### 2015-04-01 swmpr_manu.Rnw
Kahle13 ref update
updated text for function name changes - aggregate to aggreswmp and aggregate_metab to aggremetab

The analysis functions range from general purpose tools for time series analysis to more specific functions for working with continuous monitoring data in estuaries (\Cref{tab:analyze}).  The general purpose tools are \texttt{swmpr} methods that were developed for existing generic functions in the R base installation or relevant packages.  These functions include \texttt{swmpr} methods for \texttt{aggregate}, \texttt{filter}, and \texttt{approx} to deal with missing or noisy data and more general functions for exploratory data analysis, such as \texttt{plot}, \texttt{lines}, and \texttt{hist} methods.  Decomposition functions, \texttt{decomp} and \texttt{decomp\_cj}, are provided as relatively simple approaches for decomposing time series into additive or multiplicative components. Functions to estimate and plot ecosystem metabolism from combined water quality and weather data are provided by the \texttt{ecometab} and \texttt{plot\_metab} functions.  The analysis functions may or may not return a \texttt{swmpr} object depending on whether further processing with \texttt{swmpr} methods is possible from the output.

The \texttt{aggregate} function aggregates parameter data for a \texttt{swmpr} object by set units of time.  This function is most useful for aggregating noisy data to evaluate trends on longer time scales or to simply reduce the size of a dataset.  Data can be aggregated by years, quarters, months, weeks, days, or hours by a predefined function, which defaults to the mean.  A \texttt{swmpr} object is returned for the aggregated data, although the \texttt{datetimestamp} vector will be converted to a date object if the aggregation period is a day or longer.  Days are assigned to the date vector if the aggregation period is a week or longer based on the \texttt{round} method for IDate objects created in the data.table package \cite{Dowle14}.  Additionally, the method of treating \texttt{NA} values for the aggregation function should be noted since this may greatly affect the quantity of data that are returned, particularly for nutrient data.

@@

funcs <- c('aggregate.swmpr', 'aggregate\\_metab.swmpr', 'ecometab.swmpr', 'decomp.swmpr', 'decomp\\_cj.swmpr', 'hist.swmpr', 'lines.swmpr', 'na.approx.swmpr', 'plot.swmpr', 'plot\\_metab.swmpr', 'plot\\_summary.swmpr', 'smoother.swmpr')
funcs <- paste0('\\texttt{', funcs, '}')

@@

latex(
  to_tab[, 'Description', drop = F],
  file = '',
  caption = "Analysis functions available from the SWMPr package.  Full documentation for each function is in the help file (e.g., execute \\texttt{?aggregate.swmpr} at the command line).",
  rowlabel = 'Function',
  colheads = 'Description',
  rowname = to_tab$Functions,
  caption.loc = 'top',
  col.just=c("p{3.5in}"), 
  label = 'tab:analyze', 
  table.env = FALSE
  )
@

@@

# get mean DO by months
aggregate(dat, 'months', params = c('do_mgl'))
@

@@

For brevity, only the \texttt{reserve\_map} function is discussed.  This function can be used to create a map with all stations at a reserve using functions in the ggmap package \cite{ggmap13}. The current function is limited to Google maps of four types that can be set with the \texttt{map\_type} argument: terrain (default), satellite, roadmap, or hybrid.  The \texttt{zoom} argument may have to be chosen through trial and error depending on the spatial extent of the reserve.  See the help documentation for the \texttt{ggmap} function for more info on \texttt{zoom}.

>>>>

The analysis functions range from general purpose tools for time series analysis to more specific functions for working with continuous monitoring data in estuaries (\Cref{tab:analyze}).  The general purpose tools are \texttt{swmpr} methods that were developed for existing generic functions in the R base installation or relevant packages.  These functions include \texttt{swmpr} methods for \texttt{aggreswmp}, \texttt{filter}, and \texttt{approx} to deal with missing or noisy data and more general functions for exploratory data analysis, such as \texttt{plot}, \texttt{lines}, and \texttt{hist} methods.  Decomposition functions, \texttt{decomp} and \texttt{decomp\_cj}, are provided as relatively simple approaches for decomposing time series into additive or multiplicative components. Functions to estimate and plot ecosystem metabolism from combined water quality and weather data are provided by the \texttt{ecometab} and \texttt{plot\_metab} functions.  The analysis functions may or may not return a \texttt{swmpr} object depending on whether further processing with \texttt{swmpr} methods is possible from the output.

The \texttt{aggreswmp} function aggregates parameter data for a \texttt{swmpr} object by set units of time.  This function is most useful for aggregating noisy data to evaluate trends on longer time scales or to simply reduce the size of a dataset.  Data can be aggregated by years, quarters, months, weeks, days, or hours by a predefined function, which defaults to the mean.  A \texttt{swmpr} object is returned for the aggregated data, although the \texttt{datetimestamp} vector will be converted to a date object if the aggregation period is a day or longer.  Days are assigned to the date vector if the aggregation period is a week or longer based on the \texttt{round} method for IDate objects created in the data.table package \cite{Dowle14}.  Additionally, the method of treating \texttt{NA} values for the aggregation function should be noted since this may greatly affect the quantity of data that are returned, particularly for nutrient data.

@@

funcs <- c('aggreswmp.swmpr', 'aggremetab.swmpr', 'ecometab.swmpr', 'decomp.swmpr', 'decomp\\_cj.swmpr', 'hist.swmpr', 'lines.swmpr', 'na.approx.swmpr', 'plot.swmpr', 'plot\\_metab.swmpr', 'plot\\_summary.swmpr', 'smoother.swmpr')
funcs <- paste0('\\texttt{', funcs, '}')

@@

latex(
  to_tab[, 'Description', drop = F],
  file = '',
  caption = "Analysis functions available from the SWMPr package.  Full documentation for each function is in the help file (e.g., execute \\texttt{?aggreswmp.swmpr} at the command line).",
  rowlabel = 'Function',
  colheads = 'Description',
  rowname = to_tab$Functions,
  caption.loc = 'top',
  col.just=c("p{3.5in}"), 
  label = 'tab:analyze', 
  table.env = FALSE
  )
@

@@

# get mean DO by months
aggreswmp(dat, 'months', params = c('do_mgl'))
@

@@

For brevity, only the \texttt{reserve\_map} function is discussed.  This function can be used to create a map with all stations at a reserve using functions in the ggmap package \cite{Kahle13}. The current function is limited to Google maps of four types that can be set with the \texttt{map\_type} argument: terrain (default), satellite, roadmap, or hybrid.  The \texttt{zoom} argument may have to be chosen through trial and error depending on the spatial extent of the reserve.  See the help documentation for the \texttt{ggmap} function for more info on \texttt{zoom}.

### 2015-04-01 swmpr_manu.tex
Kahle13 ref update
updated text for function name changes - aggregate to aggreswmp and aggregate_metab to aggremetab

The analysis functions range from general purpose tools for time series analysis to more specific functions for working with continuous monitoring data in estuaries (\Cref{tab:analyze}).  The general purpose tools are \texttt{swmpr} methods that were developed for existing generic functions in the R base installation or relevant packages.  These functions include \texttt{swmpr} methods for \texttt{aggregate}, \texttt{filter}, and \texttt{approx} to deal with missing or noisy data and more general functions for exploratory data analysis, such as \texttt{plot}, \texttt{lines}, and \texttt{hist} methods.  Decomposition functions, \texttt{decomp} and \texttt{decomp\_cj}, are provided as relatively simple approaches for decomposing time series into additive or multiplicative components. Functions to estimate and plot ecosystem metabolism from combined water quality and weather data are provided by the \texttt{ecometab} and \texttt{plot\_metab} functions.  The analysis functions may or may not return a \texttt{swmpr} object depending on whether further processing with \texttt{swmpr} methods is possible from the output.

The \texttt{aggregate} function aggregates parameter data for a \texttt{swmpr} object by set units of time.  This function is most useful for aggregating noisy data to evaluate trends on longer time scales or to simply reduce the size of a dataset.  Data can be aggregated by years, quarters, months, weeks, days, or hours by a predefined function, which defaults to the mean.  A \texttt{swmpr} object is returned for the aggregated data, although the \texttt{datetimestamp} vector will be converted to a date object if the aggregation period is a day or longer.  Days are assigned to the date vector if the aggregation period is a week or longer based on the \texttt{round} method for IDate objects created in the data.table package \cite{Dowle14}.  Additionally, the method of treating \texttt{NA} values for the aggregation function should be noted since this may greatly affect the quantity of data that are returned, particularly for nutrient data.

% table for analysis functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Analysis functions available from the SWMPr package.  Full documentation for each function is in the help file (e.g., execute \\texttt{?aggregate.swmpr} at the command line).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), label = "tab:analyze",     table.env = FALSE)%
\begin{table}[!tbp]
\caption{Analysis functions available from the SWMPr package.  Full documentation for each function is in the help file (e.g., execute \texttt{?aggregate.swmpr} at the command line).\label{tab:analyze}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{aggregate.swmpr}&Aggregate \texttt{swmpr} objects for different time periods - years, quarters, months,  weeks, days, or hours.  The aggregation function defaults to the mean.\tabularnewline
\texttt{aggregate\_metab.swmpr}&Aggregate metabolism data from a \texttt{swmpr} object.  This is primarly used within \texttt{plot\_metab} but may be useful for simple summaries of raw metabolism data.\tabularnewline
\texttt{ecometab.swmpr}&Estimate ecosystem metabolism for a combined water quality and weather dataset using the open-water method.\tabularnewline
\texttt{decomp.swmpr}&Decompose a \texttt{swmpr} time series into trend, seasonal, and residual components.  This is a simple wrapper to \texttt{decompose} \cite{Kendall83}.  Decomposition of monthly or daily trends is possible.\tabularnewline
\texttt{decomp\_cj.swmpr}&Decompose a \texttt{swmpr} time series into grandmean, annual, seasonal, and events components.  This is a simple wrapper to \texttt{decompTs} in the wq package \cite{Jassby14}.  Only monthly decomposition is possible.\tabularnewline
\texttt{hist.swmpr}&Plot a histogram for a \texttt{swmpr} object.\tabularnewline
\texttt{lines.swmpr}&Add lines to an existing plot created with \texttt{plot.swmpr}.\tabularnewline
\texttt{na.approx.swmpr}&Linearly interpolate missing data (\texttt{NA} values) in a \texttt{swmpr} object. The maximum gap size that is interpolated is defined by the arguments.\tabularnewline
\texttt{plot.swmpr}&Plot a univariate  time series for a \texttt{swmpr} object.  The parameter name must be specified.\tabularnewline
\texttt{plot\_metab.swmpr}&Plot ecosystem metabolism estimates after running \texttt{ecometab} on a \texttt{swmpr} object.\tabularnewline
\texttt{plot\_summary.swmpr}&Create summary plots of seasonal/annual trends and anomalies for a water a single paramter of interest.\tabularnewline
\texttt{smoother.swmpr}&Smooth \texttt{swmpr} objects with a moving window average.  Window size and sides (e.g., centered) can be specified, passed to \texttt{filter}.\tabularnewline
\hline
\end{tabular}\end{center}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # get data
> data(apacpwq)
> dat <- qaqc(apacpwq)
> 
> # get mean DO by months
> aggregate(dat, 'months', params = c('do_mgl'))
\end{verbatim}
\end{kframe}
\end{knitrout}

>>>>

The analysis functions range from general purpose tools for time series analysis to more specific functions for working with continuous monitoring data in estuaries (\Cref{tab:analyze}).  The general purpose tools are \texttt{swmpr} methods that were developed for existing generic functions in the R base installation or relevant packages.  These functions include \texttt{swmpr} methods for \texttt{aggreswmp}, \texttt{filter}, and \texttt{approx} to deal with missing or noisy data and more general functions for exploratory data analysis, such as \texttt{plot}, \texttt{lines}, and \texttt{hist} methods.  Decomposition functions, \texttt{decomp} and \texttt{decomp\_cj}, are provided as relatively simple approaches for decomposing time series into additive or multiplicative components. Functions to estimate and plot ecosystem metabolism from combined water quality and weather data are provided by the \texttt{ecometab} and \texttt{plot\_metab} functions.  The analysis functions may or may not return a \texttt{swmpr} object depending on whether further processing with \texttt{swmpr} methods is possible from the output.

The \texttt{aggreswmp} function aggregates parameter data for a \texttt{swmpr} object by set units of time.  This function is most useful for aggregating noisy data to evaluate trends on longer time scales or to simply reduce the size of a dataset.  Data can be aggregated by years, quarters, months, weeks, days, or hours by a predefined function, which defaults to the mean.  A \texttt{swmpr} object is returned for the aggregated data, although the \texttt{datetimestamp} vector will be converted to a date object if the aggregation period is a day or longer.  Days are assigned to the date vector if the aggregation period is a week or longer based on the \texttt{round} method for IDate objects created in the data.table package \cite{Dowle14}.  Additionally, the method of treating \texttt{NA} values for the aggregation function should be noted since this may greatly affect the quantity of data that are returned, particularly for nutrient data.

% table for analysis functions
%latex.default(to_tab[, "Description", drop = F], file = "", caption = "Analysis functions available from the SWMPr package.  Full documentation for each function is in the help file (e.g., execute \\texttt{?aggreswmp.swmpr} at the command line).",     rowlabel = "Function", colheads = "Description", rowname = to_tab$Functions,     caption.loc = "top", col.just = c("p{3.5in}"), label = "tab:analyze",     table.env = FALSE)%
\begin{table}[!tbp]
\caption{Analysis functions available from the SWMPr package.  Full documentation for each function is in the help file (e.g., execute \texttt{?aggreswmp.swmpr} at the command line).\label{tab:analyze}} 
\begin{center}
\begin{tabular}{lp{3.5in}}
\hline\hline
\multicolumn{1}{l}{Function}&\multicolumn{1}{c}{Description}\tabularnewline
\hline
\texttt{aggreswmp.swmpr}&Aggregate \texttt{swmpr} objects for different time periods - years, quarters, months,  weeks, days, or hours.  The aggregation function defaults to the mean.\tabularnewline
\texttt{aggremetab.swmpr}&Aggregate metabolism data from a \texttt{swmpr} object.  This is primarly used within \texttt{plot\_metab} but may be useful for simple summaries of raw metabolism data.\tabularnewline
\texttt{ecometab.swmpr}&Estimate ecosystem metabolism for a combined water quality and weather dataset using the open-water method.\tabularnewline
\texttt{decomp.swmpr}&Decompose a \texttt{swmpr} time series into trend, seasonal, and residual components.  This is a simple wrapper to \texttt{decompose} \cite{Kendall83}.  Decomposition of monthly or daily trends is possible.\tabularnewline
\texttt{decomp\_cj.swmpr}&Decompose a \texttt{swmpr} time series into grandmean, annual, seasonal, and events components.  This is a simple wrapper to \texttt{decompTs} in the wq package \cite{Jassby14}.  Only monthly decomposition is possible.\tabularnewline
\texttt{hist.swmpr}&Plot a histogram for a \texttt{swmpr} object.\tabularnewline
\texttt{lines.swmpr}&Add lines to an existing plot created with \texttt{plot.swmpr}.\tabularnewline
\texttt{na.approx.swmpr}&Linearly interpolate missing data (\texttt{NA} values) in a \texttt{swmpr} object. The maximum gap size that is interpolated is defined by the arguments.\tabularnewline
\texttt{plot.swmpr}&Plot a univariate  time series for a \texttt{swmpr} object.  The parameter name must be specified.\tabularnewline
\texttt{plot\_metab.swmpr}&Plot ecosystem metabolism estimates after running \texttt{ecometab} on a \texttt{swmpr} object.\tabularnewline
\texttt{plot\_summary.swmpr}&Create summary plots of seasonal/annual trends and anomalies for a water a single paramter of interest.\tabularnewline
\texttt{smoother.swmpr}&Smooth \texttt{swmpr} objects with a moving window average.  Window size and sides (e.g., centered) can be specified, passed to \texttt{filter}.\tabularnewline
\hline
\end{tabular}\end{center}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # get data
> data(apacpwq)
> dat <- qaqc(apacpwq)
> 
> # get mean DO by months
> aggreswmp(dat, 'months', params = c('do_mgl'))
\end{verbatim}
\end{kframe}
\end{knitrout}
